---
title: "Correspondence analyis "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 7, warning = F)
library(ggplot2)
library(magrittr)
library(tidyverse)
```

```{r load-datas}
drake::loadd(catch_info)
drake::loadd(refuge_info)
drake::loadd(occasion_info)
```


Just a quick look at the differences between samplings reveals, as expected that the largest differences in catch are due to fishing gear. 
The separation is evident. 

```{r}
occasion_cov <- occasion_info %>%
  mutate(year = lubridate::year(date),
         month = lubridate::month(date), 
         season = if_else(month %in% c(2,5), "dry", "wet")) %>%
  select(occasion, year, season, month) %>%
  distinct()

catch_standard <- catch_info %>%
 # standardise replicate sampling
  group_by(refuge, occasion, gear, species) %>%
  summarise(
    across(c(no_fish, total_weight), ~round(mean(., na.rm = T)* 8)), 
    .groups = "drop")

catch_wide <- catch_standard %>%
  # Convert to a matrix
  pivot_wider(id_cols = c("refuge", "occasion", "gear"), 
              names_from = "species", 
              values_from = "no_fish") %>%
  # fill un-sampled species to zero
  mutate(across(where(is_numeric), tidyr::replace_na, replace = 0),
         across(where(is.numeric), log1p)) %>%
  left_join(select(refuge_info, refuge, category), by = "refuge") %>%
  left_join(occasion_cov, by = "occasion") %>%
  select(!starts_with("sp"), starts_with("sp")) %>%
  mutate(across(1:7, factor)) 

catch_ca <- catch_wide %>%
  FactoMineR::CA(quali.sup = 1:7, graph = FALSE)

factoextra::fviz_ca_row(catch_ca, col.row = catch_wide$gear, axes = c(1,2), geom = c("point")) +
  scale_colour_viridis_d() 

```

In order to prevent masking of other patterns due to gear, and because of the sampling irregularities, we will repeat the visualisation, but this time only looking at the gill net data. 

A new pattern emerges, and now community seems to be best explained by the fish refuge it belongs to. The signal appears to be fairly weak though and is less evident when species numbers are not log-transformed. 

```{r}
catch_gn_wide <- catch_wide %>%
  filter(gear == "GN") %>%
  select(-gear)

catch_gn_ca <- catch_gn_wide %>%
  select(-(1:6)) %>%
  FactoMineR::CA(graph = FALSE)

factoextra::fviz_ca_row(catch_gn_ca, col.row = catch_gn_wide$refuge, axes = c(1,2), geom = c("point")) +
  scale_colour_viridis_d() 
```


There doesn't appear to be any evident similarities (at least in the first two axis) among seasons within a refuge, or among refuge categories within an occasion. 

```{r, fig.height=10}
catch_gn_ca$row$coord %>%
  as.data.frame() %>%
  bind_cols(catch_gn_wide) %>%
  ggplot(aes(x = `Dim 1`, y = `Dim 2`, colour = season)) +
  geom_point() +
  facet_wrap("refuge", scales = "free") +
  scale_colour_viridis_d() +
  theme_minimal()

```

```{r, fig.height=10}

catch_gn_ca$row$coord %>%
  as.data.frame() %>%
  bind_cols(catch_gn_wide) %>%
  ggplot(aes(x = `Dim 1`, y = `Dim 2`, colour = category)) +
  geom_point() +
  facet_wrap("occasion", scales = "free") +
  scale_colour_viridis_d() +
  theme_minimal()
```
