---
title: "Alpha diversity"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 7, warning = F)
library(ggplot2)
library(magrittr)
library(tidyverse)
```

```{r load-data}
drake::loadd(catch_info)
drake::loadd(refuge_info)
drake::loadd(occasion_info)
```

Claculate shannon's diversity index

```{r}
catch_standard <- catch_info %>%
 # standardise replicate sampling
  group_by(refuge, occasion, gear, species) %>%
  summarise(
    # Mean not appropriate as some surveys came empty even when they were
    # performed. For analysis need to calculate number of samplings performed
    # and standardise using that rather than the mean
    across(c(no_fish, total_weight), ~round(mean(., na.rm = T)* 8)), 
    .groups = "drop")

catch_wide <- catch_standard %>%
  # Convert to a matrix
  pivot_wider(id_cols = c("refuge", "occasion", "gear"), 
              names_from = "species", 
              values_from = "no_fish") %>%
  # fill un-sampled species to zero
  mutate(across(where(is_numeric), tidyr::replace_na, replace = 0),
         across(where(is.numeric), log1p)) %>%
  select(!starts_with("sp"), starts_with("sp")) 

div <- catch_wide %>%
  select(starts_with("sp")) %>%
  as.matrix() %>% 
  vegan::diversity(index = "shannon", base = exp(1))

diversity <- catch_wide %>%
  select(!starts_with("sp")) %>%
  mutate(diversity_alpha = div) 
```

Having a look at the distribution we can see that the diversity of Line fishing is lower than gill net and fike nets. 
Also it seems it looks like need any special transformation for modelling; particularly if only gill net diversity values are included. 

```{r, fig.height=5}
diversity %>%
  ggplot(aes(x = diversity_alpha, fill = gear)) +
  geom_histogram(binwidth = 0.1, position = "stack", center = 0) +
  scale_fill_viridis_d() +
  facet_wrap("gear", scales = "free_y")
```

It also appears, as it was seen in previous data explorations, that the diversity at the end of the wet season seems to be the highest while the diversity at the end of the dry season or beginning of the wet season seems to be the lowest. 

```{r, fig.height=8}
diversity %>%
  ggplot(aes(x = occasion, y = diversity_alpha)) +
  geom_line(aes(group = refuge, colour = refuge), alpha = 0.25) +
  geom_boxplot() +
  scale_color_viridis_d() +
  facet_wrap("gear", ncol = 1)
```

## Covariates of diversity

We now integrate other data about sites and locations that can be useful at explaining diversity values. 
This can be later used in a statistical model. 

```{r}
occasion_cov <- occasion_info %>%
  mutate(year = lubridate::year(date),
         month = lubridate::month(date), 
         season = if_else(month %in% c(2,5), "dry", "wet"), 
         water_level = (gauge_start + gauge_finish) / 2, 
         visibility = secchi_depth, 
         aquatic_plant_index = aquatic_plant_area * aquatic_plant_density, 
         across(c("month", "year"), as.character)) %>%
  select(refuge, occasion, year, season, month, water_level, water_temp, visibility,
         aquatic_plant_area, aquatic_plant_index, brush_park) %>%
  distinct()

refuge_cov <- refuge_info %>%
  mutate(dry_area = area_dry_ha, 
         dry_depth = depth_dry_m, 
         rf_area_connected = log(rf_area_connected_in_wet_season_ha), 
         across(category, as.character)) %>%
  select(refuge, category, dry_area, dry_depth, large_water_body)

diversity_cov <- diversity %>%
  left_join(refuge_cov, by = "refuge") %>%
  left_join(occasion_cov, by = c("occasion", "refuge")) %>%
  filter(gear == "GN")
```

### Missing values

Water temperature and the aquatic_plant measurements appear to have small levels of missing values. 
We might need to check wheter there is a systematic bias on missing values before including these two variables in a model. 

```{r}
inspectdf::inspect_na(diversity_cov) %>% filter(pcnt > 0) %>% knitr::kable()
```

It appears that all cases of missing water temperature are within the first occasion at the end of the wet season in 2012. 

```{r}
na_temp_div <- filter(diversity_cov, is.na(water_temp)) %>% select(-water_temp, -gear)
no_na_temp_div <- filter(diversity_cov, !is.na(water_temp)) %>% select(-water_temp, -gear)
inspectdf::inspect_cat(na_temp_div, no_na_temp_div) %>% inspectdf::show_plot()
```

### Variable distributions

Some variables might be better behaved with a log-transformation. These include, the plant index, dry_area, dry_depth, and perhaps visibility and water level. 

```{r}
inspectdf::inspect_num(diversity_cov) %>% inspectdf::show_plot()
```

```{r}
diversity_cov %>%
  mutate(across(c(aquatic_plant_index, dry_area, dry_depth, visibility, water_level), log1p)) %>%
  inspectdf::inspect_num() %>% inspectdf::show_plot()
```

### Categotries balance

Categories appear to be well balanced except for 2012 year which would be potentially confounded with occasion 01. 

```{r}
inspectdf::inspect_cat(diversity_cov) %>% inspectdf::show_plot()
```

### Correlations

Larger refuges also tend to be deeper so maybe we can integrate these two variables into one or just use one of them as an explanatory variable. 

Also there seems to be a positive relationship between visibility, water level, aquatic plants and diversity. And a negative between those and water temperature. 

Dry depth and dry area appear to be non-seasonal factors that are positively correlated with diversity. 

These results generally agree with the multivariate analysis of [seasonality]("seasonality.md")

```{r}
inspectdf::inspect_cor(diversity_cov, method = "spearman") %>% inspectdf::show_plot()
```

When using a Pearson correlation coefficient, patterns hold except for the aquatic plant index which reverses. This shows the importance of the very large values in this index which are reduced when using the spearman coefficient. 

```{r}
inspectdf::inspect_cor(diversity_cov, method = "pearson") %>% inspectdf::show_plot()
```

### Statistical model

A quick exploration of the diversity parameters. 

First, we are interested on investigating whether there are diversity differences across seasons, and particularly the beginning and the end of each of them. 
First, we use a extra simple model with seasons/months as categorical values which is looking for significant differences among them.
Then, we look at a more mechanistic point of view which looks at the environmental factors that promote diversity.

```{r}
m_c <- lm(diversity_alpha ~ month + year, data = diversity_cov)
summary(m_c)
m_m <- lm(diversity_alpha ~ water_level + aquatic_plant_area + brush_park + water_temp, 
           data = diversity_cov)
summary(m_m)
```

We found that the average difference between the beginning and end of the seasons is fairly small and that the differences among years is also relatively small. 

The model with mechanistic variables show some significant relationships but the R2 of that model is really small. Suggesting that they are not as important on its own and account for only a small amount of variability that can be explained by treating seasonality as a categorical variable. 

If we use models like this in the actual paper we should focus on the categorical model and explain how environmental co-variates change across seasons. 


```{r}
diversity_wide <- diversity_cov %>%
  mutate(year_s = (as.numeric(substr(occasion, 3, 4))-1) %/% 4, 
         month_s = paste0("m_", month)) %>% 
  pivot_wider(id_cols = c(refuge, year), 
              names_from = month_s, 
              values_from = diversity_alpha) 

diversity_wide %>%
  inspectdf::inspect_cor() %>%
  inspectdf::show_plot()

library(brms)

m_dw <- brm(m_8 ~ m_5 + (1 | year) + (1 | refuge), data = diversity_wide)
summary(m_dw)
```

