---
title: "Seasonality"
output: github_document
---

Here we explore the seasonality of the sampling. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 7)
library(ggplot2)
library(magrittr)
```

```{r}
drake::loadd(occasion_info)
drake::loadd(refuge_info)
```

```{r season-boxplots, fig.height=5}
# Function to plot a numeric variable
boxplot_occasion <- function(var){
  occasion_info %>%
    dplyr::select(-sampling) %>%
    dplyr::distinct() %>%
    ggplot(aes_string(x = "occasion", y = var))  +
    geom_boxplot() +
    geom_point(shape = 21) +
    labs(y = snakecase::to_sentence_case(var), 
         title = snakecase::to_sentence_case(var))
}
# Loop over the numeric variables
vars <- colnames(occasion_info)[5:16]
purrr::map(vars, boxplot_occasion)
```

Let's have a closer look at the water level. But this time in logarithmic scale because there is a large variability on the rice field water levels among sites. 

```{r water-level, fig.height=11}
occasion_info %>%
  dplyr::select(-sampling) %>%
  dplyr::distinct() %>%
  dplyr::left_join(refuge_info, by = "refuge") %>%
  dplyr::mutate(occasion_numeric = as.numeric(stringr::str_extract(occasion, "[0-9]+")), 
                dry_season = (occasion_numeric %% 4) %in% c(2,3), 
                refuge = forcats::fct_reorder(refuge, as.numeric(factor(category_name)))) %>%
  ggplot(aes(x = occasion, y = rf_water_level, group = refuge, colour = category_name)) +
  geom_point(aes(colour = category_name)) +
  geom_line() +
  # scale_colour_manual(values = c("black", "red")) +
  scale_y_continuous(trans = "log1p", breaks = c(1,10,100)) +
  facet_wrap("refuge", ncol = 4) +
  theme(legend.position = "top", legend.direction = "vertical", 
        axis.text.x = element_text(angle = 90))
```

We can see there is a lot of variation in the flooding patterns. Many sites are flodded in both of the we season occasions consistently throughout the years, but many miss the flood at least once or tice. It's apparent that including this factor somehow in the analysis might become important when comparing and accounting for biodiversity in the wet season in particular. 

Flooding patterns do not seem to be at all related to the site categories. 

```{r pca-water-level}
ri <- refuge_info %>%
  dplyr::select(refuge, province, category_name, large_water_body, type_inlet_outlet, 
                depth_wet_m, depth_dry_m, drought_freq, flooding_freq, rf_area_connected_in_dry_season_ha, 
                rf_area_connected_in_wet_season_ha)

pca_analysis <- occasion_info %>%
  dplyr::select(-sampling) %>%
  dplyr::distinct() %>%
  dplyr::select(refuge, occasion, rf_water_level) %>%
  dplyr::mutate(rf_water_level = log(rf_water_level + 1)) %>%
  tidyr::pivot_wider(id_cols = "refuge", values_from = "rf_water_level", names_from = "occasion") %>% 
  dplyr::mutate(dplyr::across(dplyr::everything(), tidyr::replace_na, 0)) %>% 
  dplyr::left_join(ri, , by = "refuge") %>% 
  as.data.frame() %>%
  magrittr::set_rownames(.$refuge) %>%
  dplyr::select(-refuge) %>% 
  FactoMineR::PCA(quali.sup = 14:17, quanti.sup = 18:23, graph = F)

factoextra::fviz_pca_ind(pca_analysis, habillage = "category_name", addEllipses = T)
```

