---
title: "Seasonality"
output: github_document
---

Here we explore the seasonality of the sampling. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 7)
library(ggplot2)
library(magrittr)
library(tidyverse)
```

```{r}
drake::loadd(occasion_info)
drake::loadd(refuge_info)
```

```{r season-boxplots, fig.height=5}
# Function to plot a numeric variable
boxplot_occasion <- function(var){
  occasion_info %>%
    dplyr::select(-sampling) %>%
    dplyr::distinct() %>%
    ggplot(aes_string(x = "occasion", y = var))  +
    geom_boxplot() +
    geom_point(shape = 21) +
    labs(y = snakecase::to_sentence_case(var), 
         title = snakecase::to_sentence_case(var))
}
# Loop over the numeric variables
vars <- colnames(occasion_info)[5:16]
purrr::map(vars, boxplot_occasion)
```

Let's have a closer look at the water level. But this time in logarithmic scale because there is a large variability on the rice field water levels among sites. 

```{r water-level, fig.height=11}
occasion_info %>%
  dplyr::select(-sampling) %>%
  dplyr::distinct() %>%
  dplyr::left_join(refuge_info, by = "refuge") %>%
  dplyr::mutate(occasion_numeric = as.numeric(stringr::str_extract(occasion, "[0-9]+")), 
                dry_season = (occasion_numeric %% 4) %in% c(2,3), 
                refuge = forcats::fct_reorder(refuge, as.numeric(factor(category_name)))) %>%
  ggplot(aes(x = occasion, y = rf_water_level, group = refuge, colour = category_name)) +
  geom_point(aes(colour = category_name)) +
  geom_line() +
  # scale_colour_manual(values = c("black", "red")) +
  scale_y_continuous(trans = "log1p", breaks = c(1,10,100)) +
  facet_wrap("refuge", ncol = 4) +
  theme(legend.position = "top", legend.direction = "vertical", 
        axis.text.x = element_text(angle = 90))
```

We can see there is a lot of variation in the flooding patterns. Many sites are flodded in both of the we season occasions consistently throughout the years, but many miss the flood at least once or tice. It's apparent that including this factor somehow in the analysis might become important when comparing and accounting for biodiversity in the wet season in particular. 

Flooding patterns do not seem to be at all related to the site categories. 

```{r pca-water-level}
ri <- refuge_info %>%
  dplyr::select(refuge, province, category_name, large_water_body, type_inlet_outlet, 
                depth_wet_m, depth_dry_m, drought_freq, flooding_freq, rf_area_connected_in_dry_season_ha, 
                rf_area_connected_in_wet_season_ha)

pca_analysis <- occasion_info %>%
  dplyr::select(-sampling) %>%
  dplyr::distinct() %>%
  dplyr::select(refuge, occasion, rf_water_level) %>%
  dplyr::mutate(rf_water_level = log(rf_water_level + 1)) %>%
  tidyr::pivot_wider(id_cols = "refuge", values_from = "rf_water_level", names_from = "occasion") %>% 
  dplyr::mutate(dplyr::across(where(is.numeric),  tidyr::replace_na, 0)) %>% 
  dplyr::left_join(ri) %>% 
  #as.data.frame() %>%
  #magrittr::set_rownames(.$refuge) %>%
  dplyr::select(-refuge) %>% 
  FactoMineR::PCA(quali.sup = 14:17, quanti.sup = 18:23, graph = F)

factoextra::fviz_pca_ind(pca_analysis, habillage = "category_name", addEllipses = T)
```

## Seasonality coovariates

Now we explore what characterises each sampling occasion to better understand that. 

We can decompose the seasonality variables into several axis. 

The firs axis is related to water level: gauge start, gauge finish, secci depth, and rice field water deptth are highly correlated. This shows a clear but still muddy separation across months, and seasons. 

The second group is related to nutrients: aquatic plant area, aquatic plant density, conductivity and phosphate point in the same direction. Interestingly, it is negatively correlated to nitrogen (albeit not by much).

The third component is related to more brush park and less water temperature. Second and third dimension tend to be more related to year. It appears that, brush park increased across years while temperature decreased. Patterns are not very clear but it shows that including these variables as main effects might not account for the whole variablity. 

```{r}
season_pca <- occasion_info %>%
  mutate(year = lubridate::year(date),
         month = lubridate::month(date), 
         season = if_else(month %in% c(2,5), "dry", "wet"), 
         across(c("month", "year"), as.character)) %>%
  select(occasion, year, month, season, everything(), -sampling, -weather, -date, - refuge, -illegal_fishing, -water_bird, -other_animal, -total_hours) %>% 
  distinct() %>%
  # mutate(across(c(illegal_fishing, water_bird, other_animal), as.numeric)) %>%
  FactoMineR::PCA(quali.sup = 1:4)

factoextra::fviz_pca_var(season_pca)
factoextra::fviz_pca_var(season_pca, axes = 3:4)
factoextra::fviz_pca_ind(season_pca, geom = "point", habillage = "season", addEllipses = T)
factoextra::fviz_pca_ind(season_pca, geom = "point", habillage = "month", addEllipses = T)
factoextra::fviz_pca_ind(season_pca, geom = "point", habillage = "occasion", addEllipses = T, axes = 3:4)
factoextra::fviz_pca_ind(season_pca, geom = "", habillage = "year", addEllipses = T, axes = 3:4)
FactoInvestigate::Investigate(season_pca)
```

Looks like we can focus on a few variables only: aquatic plant area, gauge_start, water_temp, phosphate and brush_park

## Now comparing variables from the dry season alone

```{r}
oc <- occasion_info %>%
  mutate(year = lubridate::year(date),
         month = lubridate::month(date), 
         season = if_else(month %in% c(2,5), "dry", "wet"), 
         across(c("month", "year"), as.character)) %>%
  select(occasion, year, month, season, aquatic_plant_area, gauge_start, water_temp, phosphate, brush_park, refuge) %>%
  distinct() %>%
  filter(season == "dry")

oc_m2 <- oc %>%
  filter(month == "2")

oc_m5 <- oc %>%
  filter(month == "5")

ocd <- full_join(oc_m2, oc_m5, by = c("refuge", "year"), suffix = c("_m2", "_m5"))

dry_s_pca <- ocd %>%
  select(where(is.numeric)) %>%
  FactoMineR::PCA()

# FactoInvestigate::Investigate(dry_s_pca)
```

There is a strong correlation between the occasion variables at the beginning and the end of the season so no need to use both. There seem to be a bit of a difference between gauge start and water temp so these could be included as the difference as well. 
